---
title: "covid 19"
author: "Deniss Gonzalez Ramirez"
date: "November 18, 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata$state<-as.factor(kRdata$state)
kRdata$result<-as.factor(kRdata$result)
kRdata$sex<-as.factor(kRdata$sex)
kRdata$Sex<-as.factor(kRdata$Sex)

```


### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
barchartGC(~result + Sex,data=kRdata)
```


The bar chart above depicts that women are alive more than men, with there being more women than men. Men ultimately had more deaths than women regardless of the amount of women and men that were a part of the study.




```{r}
barchartGC(~result + Sex,data=kRdata, type="percent")

```


In terms of percentages of women and men who were alive and dead the bar plot above shows that the relationship is reversed, meaning that on the alive side more percentage of females were alive and and less were dead. Looking at the data we can conclude that something is going on because if nothing were, the percentages would be equal on both sides.


  

  
## Numerical results


```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The first table above shows that the alive group was 56% female and 44% male. In the dead group females accounted for 36% and the males accounted for 64%. Based on the information above we can conclude that there is a backwards relationship between alive and dead and the males and females that were a part of the study. Females were the predominantly alive group and males comprised most of the dead group.


In the second table 1.4% of the females ended up dead whereas 3% of the males in the study ended up dead. The percentages of males that died is more than double the percent of females that died.


## Inferential results 


```{r}
chisq.test(table1)
chisqtestGC(table1)
```

The Chi-squared test above shows that 1831 were alive and 1411 males were alive. The females dead was 25 and 44 males were dead. The odds of a female dying is (25/1831) and the odds of a male dying is (1411/44). The odds ratio of a male dying to a female is (44/1411)/(25/1831) which is 2.3.


```{r}
fisher.test(table1)
```


The Fisher test gives us the odds of dying for males is (44/1411) and the odds of females dying is (25/1831). When  you divide these two numbers you get 2.3; the odds ratio is 2.3 .This shows that males are 2.3 times more likely to die. The confidence interval are on the low end at 1.4 to 3.9 being the high end.


## Conclusions



In conclusionthere is a significant statistical difference with men and women. It is evident that more men die from COVID in comparison to women. The p-value is very small so we can reject the null hypothesis. More importantly we can see that more men die from this disease than women. 
